---
title: "Data Vis & Models"
author: "Ranqing Lan - updated Paul Bloom"
date: "Created 2023-04-21, updated 2024-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(lme4)
library(Matrix)
library(lmerTest)
library(brms)
library(rstan)
library(emmeans)
library(tidyverse)
library(readxl,quietly=TRUE,warn.conflicts=FALSE)
library(lubridate,quietly=TRUE,warn.conflicts=FALSE )
library(zoo)
library(psych)
library(openxlsx)
library(corrplot)
library(sjPlot)
library(ggplot2)
options(scipen = 999)
library(purrr)
library(broom)
library(broom.mixed)

# function for getting odds ratios dataframe from brms logistic regression
get_or_brms = function(model){
  out_df = fixef(model) %>%
    exp() %>%
    as.data.frame() %>%
    mutate(term = row.names(.))
  return(out_df)
}
```

## Explanatory Analyses {.tabset}

### Setup

```{r, include = FALSE}
df <- read_csv('cleaned_data/weekly_data_with_stb.csv')

b= df %>%
  group_by(ID, home_min_idmean, travel_km_idmean, entropy_idmean) %>%
  count()
  
## Standardize geolocation features
df <- df %>%
  group_by(ID) %>%
  mutate(
    entropy_dev = scale(entropy_dev, center = FALSE, scale = TRUE),
    home_h_dev = scale(home_h_dev, center = FALSE, scale = TRUE),
    travel_km_dev = scale(travel_km_dev, center = FALSE, scale = TRUE),
    entropy_dev_lead1week = scale(entropy_dev_lead1week, center = FALSE, scale = TRUE),
    home_h_dev_lead1week = scale(home_h_dev_lead1week, center = FALSE, scale = TRUE),
    travel_km_dev_lead1week = scale(travel_km_dev_lead1week, center = FALSE, scale = TRUE)) %>%
  ungroup()

# Bring in data for exclusions based on home distance
homedist = read.csv('/Volumes/columbia/MAPS_Data/Analysis/Wrangling/home_location/maps_home_distances_from_nyc_pitt.csv')
df = left_join(df, dplyr::select(homedist, ID=subject_id, min_study_site_distance), by = 'ID')
df = mutate(df, homedist_exclude = min_study_site_distance > 100)

# join in self-report data
load('/Volumes/columbia/MAPS_Data/Analysis/Wrangling/Self_Report/data/self_report_02_07_2024.Rda')
selfreport_v8 = mutate(selfreport_v8, SSI19_baseline_imp1month = coalesce(SSI19_i, SSI19_1),
                       SSI19_baseline_imp1month = scale(SSI19_baseline_imp1month),
                       Control = ifelse(GROUP == 'Control',1, 0))
df = left_join(df, selfreport_v8, by = 'ID')

# remove weeks with fewer than 3 days with GPS
sum(df$gps_missing3)

# scale btw-person variables
df = mutate(df,
            entropy_idmean = scale(round(entropy_idmean, 3)),
            travel_km_idmean = scale(round(travel_km_idmean, 3)),
            home_h_idmean = scale(round(home_min_idmean/60, 3)))

df = dplyr::mutate(df, 
                   event_cat1_lead1week = ifelse(ID==2013 & week_start_wednesday== as.Date('2019-09-18'),
                                                 0, event_cat1_lead1week),
                   event_cat1 = ifelse(ID==2013 & week_start_wednesday== as.Date('2019-09-25'),
                                                 0, event_cat1),
                   event_cat1_lead2week = ifelse(ID==2013 & week_start_wednesday== as.Date('2019-09-11'),
                                                 0, event_cat1_lead2week),
                   event_cat1_lead3week = ifelse(ID==2013 & week_start_wednesday== as.Date('2019-09-04'),
                                                 0, event_cat1_lead3week))

length(unique(df$ID))
sum(df$ema_trigger, na.rm = TRUE)
sum(df$event_cat1)
```

#### Checks for data availability 
```{r}
b = df %>%
  group_by(ID) %>%
  summarise(n_obs_home = sum(!is.na(home_h_dev)),
            n_obs_travel= sum(!is.na(travel_km_dev)),
            n_obs_entropy = sum(!is.na(entropy_dev)),
            ssi_i = sum(!is.na(SSI19_i)),
            ssi_1 = sum(!is.na(SSI19_1)),
            ssi_best = sum(!is.na(SSI19_baseline_imp1month)))

c = df %>%
  dplyr::filter(is.na(SSI19_i))

e = selfreport_v8 %>%
  dplyr::filter(is.na(SSI19_i)) %>%
  dplyr::select(ID, SITE, GROUP, contains('SSI'), contains('MFQ'))
```

### EMA Triggers (Ideation)

```{r}
fit_ema_concurrent_frequentist = glmer(ema_trigger  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID) , data = df,
              family = binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

fit_ema_concurrent = brm(ema_trigger  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID), data = df,
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_ema_lead1week = glmer(ema_trigger_lead1week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                            week_num + (1|ID) , data = df,
              family = binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))


# Models asking whether triggers are associated with changes in geolocation variables the following week 
fit_homestay_lead1_week = lmer(home_h_dev_lead1week  ~ ema_trigger + schoolclosure + week_num + (1|ID) , data = df)

fit_entropy_lead1_week = lmer(entropy_dev_lead1week  ~ ema_trigger + schoolclosure + week_num + (1|ID) , data = df)

fit_travel_km_lead1_week = lmer(travel_km_dev_lead1week  ~ ema_trigger + schoolclosure + week_num + (1|ID) , data = df)


tab_model(fit_homestay_lead1_week, fit_entropy_lead1_week, fit_travel_km_lead1_week)
```

#### Tables of stats

```{r}
sjPlot::tab_model(fit_ema_concurrent_frequentist, fit_ema_lead1week)

summary(fit_ema_concurrent_frequentist)
```

#### Plots odds ratios
```{r}
df_or_ema_within <- broom::tidy(fit_ema_lead1week, conf.int=TRUE, conf.level = 0.95, exponentiate=TRUE)  %>%
  filter(term %in% c("home_h_dev", "entropy_dev", "travel_km_dev")) %>%
  dplyr:: select(term, estimate, std.error, p.value, conf.low, conf.high) %>% 
  mutate(., sig = ifelse((conf.low > 1 & conf.high > 1) |
                                            (conf.low < 1 & conf.high < 1), '*', ''))


colors = c( "lightblue", "red","black")
or_ema_within_plot <- df_or_ema_within %>%
    mutate(term = dplyr::recode(term, 'home_h_dev' ='Home Stay', 'entropy_dev'='Entropy', "travel_km_dev" = "Distance Traveled")) %>%
    ggplot(data=., aes(x = term, y = estimate)) +
    geom_point(position = position_dodge(0.6), aes(color = term)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = term),
                    lwd = 1, width = 0, position = position_dodge(0.6)) +
      labs(x = NULL, y = "Adjusted Odds Ratio", color ="black", title = 'Within-Person Suicidal Ideation Models') +
      geom_hline(yintercept = 1, lty = 2) +
      geom_text(aes(label = sig, y = conf.high*1.1), size = 5) +
      labs(color = NULL) +theme_bw()  +
      scale_color_manual(values = colors) +
      theme(strip.text.y = element_blank(),
            axis.text.x = element_text(colour = "black"),
          axis.text.y = element_text(colour = "black"),
          legend.position = 'none')  +
      scale_y_continuous(limits = c(0.25,4.2), breaks=seq(0.25,4,0.75)) +
      coord_flip()

or_ema_within_plot
```

### Suicide Events  

```{r}
fit_event_concurrent = brm(event_cat1  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID), data = df,
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_event_concurrent_frequentist = glmer(event_cat1  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID), data = df,
              family = binomial(link='logit'), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

df$month = format(df$week_start_wednesday, "%m")
df = mutate(df, summerbreak = ifelse(month %in% c('07', '09'), 1, 0))

fit_event_lead1week = brm(event_cat1_lead1week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID) , data = df,
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_event_lead1week_frequentist = glmer(event_cat1_lead1week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID) , data = df,
              family = binomial(link='logit'), control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

fit_event_lead2week = brm(event_cat1_lead2week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID) , data = df,
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_event_lead3week = brm(event_cat1_lead3week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID) , data = df,
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_event_lead1week_homestay_only = brm(event_cat1_lead1week  ~  home_h_dev + home_h_idmean + week_num + (1|ID) , data = df,
                                        family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

# Sensitivity check excluding participants with home distance over 100 miles from a study site
fit_event_concurrent_homedist_exclude = brm(event_cat1  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID),
                data = df %>% dplyr::filter(., homedist_exclude == 0),
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)

fit_event_lead1week_homedist_exclude = brm(event_cat1_lead1week  ~  home_h_dev + entropy_dev + travel_km_dev + 
                                           home_h_idmean + entropy_idmean + travel_km_idmean + 
                             week_num + (1|ID), 
                             data = df %>% dplyr::filter(., homedist_exclude == 0),
              family = bernoulli(link='logit'), iter = 2000, cores = 4, seed = 11291993)
```


#### Results table: concurrent GPS & suicide events


```{r}
sjPlot::tab_model(fit_event_concurrent, fit_event_concurrent_homedist_exclude)
```

#### Results table: GPS & next-week suicide events


```{r}
sjPlot::tab_model(fit_event_lead1week, fit_event_lead1week_homedist_exclude)
```

#### Lead 1 week table for paper

```{r}
event_model_table = broom::tidy(fit_event_lead1week_frequentist, exponentiate=TRUE, conf.int=TRUE) %>%
  dplyr::select(term, aOR=estimate, conf.low, conf.high, p.value) %>%
  dplyr::filter(term!= 'sd__(Intercept)') %>%
  mutate(aOR = round(aOR, 2), 
         `95% CI`=paste0(round(conf.low,2), ', ', round(conf.high, 2)),
         p.value = round(p.value, 3)) %>%
  dplyr::select(term, aOR, `95% CI`, p.value) %>%
  mutate(outcome = 'Suicide Events')

event_model_table_betas = broom::tidy(fit_event_lead1week_frequentist) %>%
  dplyr::select(term, beta=estimate, std.error) %>%
  dplyr::filter(term!= 'sd__(Intercept)') %>%
  mutate(beta = round(beta, 2), 
         std.error = round(std.error, 3)) %>%
  mutate(outcome = 'Suicide Events')

ema_trigger_table_betas = broom::tidy(fit_ema_lead1week) %>%
  dplyr::select(term, beta=estimate, std.error) %>%
  dplyr::filter(term!= 'sd__(Intercept)') %>%
  mutate(beta = round(beta, 2), 
         std.error = round(std.error, 3)) %>%
  mutate(outcome = 'EMA Trigger')

ema_trigger_model_table = broom::tidy(fit_ema_lead1week, exponentiate=TRUE, conf.int=TRUE) %>%
  dplyr::select(term, aOR=estimate, conf.low, conf.high, p.value) %>%
  dplyr::filter(term!= 'sd__(Intercept)') %>%
  mutate(aOR = round(aOR, 2), 
         `95% CI`=paste0(round(conf.low,2), ', ', round(conf.high, 2)),
         p.value = round(p.value, 3)) %>%
  dplyr::select(term, aOR, `95% CI`, p.value) %>%
  mutate(outcome = 'EMA Trigger')


within_participant_model_table_beta= rbind(event_model_table_betas, ema_trigger_table_betas)
within_participant_model_table = rbind(event_model_table, ema_trigger_model_table)

within_participant_model_table = left_join(within_participant_model_table_beta,
                                           within_participant_model_table,
                                           by = c('term', 'outcome')) %>%
  dplyr::select(., outcome, term, beta, std.error, aOR, `95% CI`, p.value)

write.csv(within_participant_model_table, file = 'tables/within_participant_explanatory_model_table_event_sensitivity.csv', 
          row.names = FALSE)
```

#### OR Plot for events
```{r}
df_or_cat1_within <- broom::tidy(fit_event_lead1week_frequentist, conf.int=TRUE, conf.level = 0.95, exponentiate=TRUE)  %>%
  filter(term %in% c("home_h_dev", "entropy_dev", "travel_km_dev")) %>%
  dplyr:: select(term, estimate, std.error, p.value, conf.low, conf.high) %>% 
  mutate(., sig = ifelse((conf.low > 1 & conf.high > 1) |
                                            (conf.low < 1 & conf.high < 1), '*', ''))

colors = c("lightblue", "red","black")
or_cat1_within_plot <- df_or_cat1_within %>%
    mutate(term = dplyr::recode(term, 'home_h_dev' ='Home Stay', 'entropy_dev'='Entropy', "travel_km_dev" = "Distance Traveled")) %>%
    ggplot(data=., aes(x = term, y = estimate)) +
    geom_point(position = position_dodge(0.6), aes(color = term)) +
      geom_errorbar(aes(ymin = conf.low, ymax = conf.high, color = term),
                    lwd = 1, width = 0, position = position_dodge(0.6)) +
      labs(x = NULL, y = "Adjusted Odds Ratio", title = 'Within-Person Suicidal Event Models') +
      geom_hline(yintercept = 1, lty = 2) +
      geom_text(aes(label = sig, y = conf.high*1.1), size = 5) +
      labs(color = NULL) +theme_bw()  +
      scale_color_manual(values = colors) +
      theme(strip.text.y = element_blank(),
            axis.text.x = element_text(colour = "black"),
          axis.text.y = element_text(colour = "black"),
          legend.position = 'none')  +
      scale_y_continuous(limits = c(0.25,4.2), breaks=seq(0.25,4,0.75)) +
      coord_flip() 

or_cat1_within_plot
```


```{r}
prob_df_cat1_within <- fit_event_lead1week_frequentist %>%
  emmeans::emmeans(specs = 'home_h_dev', type = 'response', at = list(home_h_dev = seq(-2, 2,.1 ))) %>% summary()

prob_plot_cat1_within <- prob_df_cat1_within %>%
  ggplot(., aes(x = home_h_dev, y = 100 * prob)) + 
  geom_line( size = 1, color = "black")  +
  geom_ribbon(aes(ymin = 100 * asymp.LCL, ymax = 100 * asymp.UCL),
              alpha=0.2,
              fill = "black") + 
  labs(x = "Home Stay (Within-participant mean-centered)", 
       y = "Probability [%] of Suicide Event\nDuring Given Week",
       title = 'Weekly Homestay & Suicidal Events') +
  labs(color = "black") +theme_bw()

prob_plot_cat1_within
```
### Between-person Analyses


```{r}
btw_summary = df %>%
  mutate(entropy_idmean = round(entropy_idmean, 2),
            travel_km_idmean = round(travel_km_idmean, 2),
            home_h_idmean = round(home_h_idmean, 2)) %>%
  dplyr::select(ID, contains('idmean'), contains('idmean'), event_cat1, ema_trigger, Control,
                SITE, age, sex, iPhone, SSI19_baseline_imp1month, GROUP) %>%
  group_by(ID, entropy_idmean, travel_km_idmean, home_h_idmean, Control, GROUP,
           SITE, age, sex, iPhone, SSI19_baseline_imp1month) %>%
  summarise(event_cat1_participant = ifelse(sum(event_cat1)>0, 1, 0),
            ema_trigger_participant = ifelse(sum(ema_trigger, na.rm=TRUE)>0, 1, 0)) %>%
  mutate(stb_pos = ifelse(Control == '1', '0', '1')) 


participant_info = selfreport_v8 %>% dplyr::filter(ID %in% df$ID) %>%
  mutate(ssi_significant_baseline = SSI19_i >= 4)

table(participant_info$ssi_significant_baseline)
table(participant_info$ssi_significant_baseline) / nrow(participant_info)



sum(is.na(participant_info$SelfReportDate_1)) 
sum(is.na(participant_info$SelfReportDate_3)) 
sum(is.na(participant_info$SelfReportDate_6))

1 - (sum(is.na(participant_info$SelfReportDate_1)) / nrow(participant_info))
1 - (sum(is.na(participant_info$SelfReportDate_3)) / nrow(participant_info))
1 - (sum(is.na(participant_info$SelfReportDate_6)) / nrow(participant_info))


baseline_si_btw_models = btw_summary %>%
  ungroup() %>%
  mutate(SSI19_baseline_imp1month_z = scale(SSI19_baseline_imp1month)) %>%
  pivot_longer(c(entropy_idmean, travel_km_idmean, home_h_idmean)) %>%
  group_by(name) %>%
  mutate(value = scale(value)) %>%
  nest() %>%
  mutate(m = purrr::map(data, ~lm(data = ., value ~ SSI19_baseline_imp1month_z + SITE + sex + iPhone + age)),
         coef = purrr::map(m, ~broom::tidy(., conf.int=TRUE))) %>%
  unnest(coef) %>%
  dplyr::select(-data, -m) %>%
  dplyr::filter(term != '(Intercept)') %>%
  mutate(type = 'Baseline\nSuicidal Ideation')

baseline_group_btw_models = btw_summary %>%
  pivot_longer(c(entropy_idmean, travel_km_idmean, home_h_idmean)) %>%
  group_by(name) %>%
  mutate(value = scale(value)) %>%
  nest() %>%
  mutate(m = purrr::map(data, ~lm(data = ., value ~ stb_pos + SITE + sex + iPhone + age)),
         coef = purrr::map(m, ~broom::tidy(., conf.int=TRUE))) %>%
  unnest(coef) %>%
  dplyr::select(-data, -m) %>%
  dplyr::filter(term != '(Intercept)') %>%
  mutate(type = 'STB History+ vs. STB History-')

btw_models_summary = rbind(baseline_si_btw_models) %>%
  dplyr::filter(grepl('_participant', term) | term == 'SSI19_baseline_imp1month_z' | term == 'stb_pos1') %>%
  mutate(name = dplyr::recode(name, 'home_h_idmean' ='Home Stay', 'entropy_idmean'='Entropy', "travel_km_idmean" = "Distance Traveled"))

colors = c( "lightblue", "red","black")


btw_models_summary = mutate(btw_models_summary, sig = ifelse((conf.low > 0 & conf.high > 0) |
                                                             (conf.low < 0 & conf.high < 0), '*', ''))

between_models_plot = ggplot(btw_models_summary, aes(x = name, y = estimate, color = name)) +
  geom_point(size = 3) + 
  geom_text(aes(label = sig, y = conf.high*1.1), size = 5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, lwd = 1) +
  facet_grid(~type) +
  geom_hline(yintercept = 0, lty = 2) +
  coord_flip() +
  theme_bw() +
  labs(y = 'Beta Estimate', x = NULL, title = '') +
  scale_color_manual(values = colors, 
                     limits = c('Distance Traveled', 'Entropy', 'Home Stay')) +
  theme(legend.position = 'none',
        axis.text.y = element_blank(),
        panel.spacing = unit(0, "lines"))


concurrent_between= rbind(get_or_brms(fit_ema_concurrent) %>% mutate(type = 'Suicidal Ideation\nESM'),
                          get_or_brms(fit_event_concurrent) %>% mutate(type = 'Suicidal Events')) %>%
  mutate(., sig = ifelse((Q2.5 > 1 & Q97.5 > 1) |
                                                             (Q2.5 < 1 & Q97.5 < 1), '*', '')) %>%
  mutate(term = dplyr::recode(term, 'travel_km_idmean'='Distance Traveled', 'home_h_idmean'='Home Stay', 'entropy_idmean'='Entropy'))




concurrent_between_models_plot = concurrent_between %>% dplyr::filter(term %in% c('Distance Traveled', 'Entropy', 'Home Stay')) %>%
  ggplot(., aes(x = term, y = Estimate, color = term)) +
    geom_point(size = 3) + 
    geom_text(aes(label = sig, y = Q97.5*1.1), size = 5) +
    geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0, lwd = 1) +
    facet_grid(~type) +
    geom_hline(yintercept = 1, lty = 2) +
    coord_flip() +
    theme_bw() +
    labs(y = 'Adjusted Odds Ratio', x = NULL, title = 'Between-Person Differences') +
    scale_color_manual(values = colors, 
                       limits = c('Distance Traveled', 'Entropy', 'Home Stay')) +
    theme(legend.position = 'none',
          axis.text.y = element_text(angle = 30, vjust = 0.5, hjust=1),
          panel.spacing = unit(0, "lines")) +
    scale_y_log10(breaks = c(.1, .3, 1, 3))


btw_grid = cowplot::plot_grid(concurrent_between_models_plot, between_models_plot, rel_widths = c(2.1, 1),
                              align = 'h', axis = 'bt')

between_models_plot

table(btw_summary$event_cat1_participant)
table(btw_summary$ema_trigger_participant)
38 + 149
173 + 14

btw_models_summary
```

#### Make Figure 1

```{r}
fig1_grid = cowplot::plot_grid(or_cat1_within_plot,
                               prob_plot_cat1_within, 
                               or_ema_within_plot,
                               btw_grid,
                               labels = c('A', 'B', 'C', 'D'))

cowplot::save_plot(fig1_grid, file = 'plots/fig1_grid_event_sensitivity.png', base_width = 10, base_height = 6)
```

### Plots/Tables for Analyses on Timing of Suicidal Events Relative to Geolocation

```{r}
# Models asking whether suicide events are associated with changes in geolocation variables the following week 
event_fit_homestay_lead1_week = lmer(home_h_dev_lead1week  ~ event_cat1 + schoolclosure + week_num + (1|ID) , data = df)

event_fit_entropy_lead1_week = lmer(entropy_dev_lead1week  ~ event_cat1 + schoolclosure + week_num + (1|ID) , data = df)

event_fit_travel_km_lead1_week = lmer(travel_km_dev_lead1week  ~ event_cat1 +schoolclosure + week_num + (1|ID) , data = df)

tab_model(event_fit_homestay_lead1_week, event_fit_entropy_lead1_week, event_fit_travel_km_lead1_week)
```

```{r}
event_predict_geo_emmgrid = rbind(
  emmeans::emmeans(event_fit_homestay_lead1_week, specs = ~event_cat1) %>% data.frame() %>% mutate(outcome = 'Home Stay'),
  emmeans::emmeans(event_fit_entropy_lead1_week, specs = ~event_cat1) %>% data.frame() %>% mutate(outcome = 'Entropy'),
  emmeans::emmeans(event_fit_travel_km_lead1_week, specs = ~event_cat1) %>% data.frame() %>% mutate(outcome = 'Distance Traveled')) %>%
  dplyr::mutate(event_recat = ifelse(event_cat1 == 0, 'No Suicide Event Last Week', 'Suicide Event Last Week'))


event_predict_geolocation_plot = ggplot(event_predict_geo_emmgrid, aes(x = event_recat, y = emmean, color = outcome)) +
  geom_point(position = position_dodge(0.1)) +
  geom_line(aes(group = outcome), position = position_dodge(0.1)) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0, position = position_dodge(0.1)) +
  scale_color_manual(values = colors) +
  theme_bw() +
  labs (y = 'Geolocation Variable Estimate\nWithin-participant centered', x = NULL, 
        title = 'Geolocation Variables as a Function of Last-Week Suicidal Events')


tidy_event_predict_geo = rbind(broom.mixed::tidy(event_fit_homestay_lead1_week, conf.int=TRUE) %>% mutate(outcome = 'Home Stay'),
                               broom.mixed::tidy(event_fit_entropy_lead1_week, conf.int=TRUE) %>% mutate(outcome = 'Entropy'),
                               broom.mixed::tidy(event_fit_travel_km_lead1_week, conf.int=TRUE) %>% mutate(outcome = 'Distance Traveled')) %>%
  dplyr::filter(term == 'event_cat1') %>%
  mutate(sig_star = ifelse(
    (conf.low > 0 & conf.high > 0) | (conf.low <0 & conf.high < 0), '*', ''))

event_predict_geolocation_plot2 = ggplot(tidy_event_predict_geo, aes(x = outcome, y = estimate, color = outcome)) +
  geom_hline(yintercept =0, lty = 2) + 
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0, lwd = 1) +
  geom_text(aes(x = outcome, y = conf.high*1.04, label = sig_star)) +
  theme_bw() +
  labs(x = 'Geolocation Metric (Within-Person Centered)', y = 'Beta Estimate (95% CI)',
       title = 'Geolocation Variables as a Function of\nLast-Week Suicidal Events') +
  scale_color_manual(values = colors) +
  theme(legend.position = 'none')




get_or_brms(fit_event_concurrent)


tidy_weeks = rbind(get_or_brms(fit_event_concurrent) %>% mutate(timing = 'Same Week'),
                   get_or_brms(fit_event_lead1week) %>% mutate(timing = 'T - 1 Week'),
                   get_or_brms(fit_event_lead2week) %>% mutate(timing = 'T - 2 Weeks'),
                   get_or_brms(fit_event_lead3week) %>% mutate(timing = 'T - 3 Weeks')) %>%
  dplyr::filter(term == 'home_h_dev') %>%
  mutate(timing = factor(timing, levels = c('T - 3 Weeks', 'T - 2 Weeks', 'T - 1 Week', 'Same Week'))) %>%
  mutate(sig_star = ifelse(
    (Q2.5 > 1 & Q97.5 > 1) | (Q2.5 > 1 & Q97.5 > 1), '*', ''))

colors = c( "lightblue", "red","black")
week_timeline_plot = ggplot(tidy_weeks, aes(x = timing, y = Estimate)) +
  geom_hline(yintercept = 1, lty = 2) + 
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = Q2.5, ymax = Q97.5), width = 0, lwd = 1) +
  geom_text(aes(x = timing, y = Q97.5*1.04, label = sig_star)) +
  scale_y_log10() +
  theme_bw() +
  labs(x = 'Time Window of Within-Participant Homestay Change\nRelative to Suicidal Events', y = 'aOR (95% CI)',
       title = 'Suicidal Events as a Function of\nTime-Windowed Homestay')


event_timing_plot = ggplot(data = df %>%
                             group_by(ID) %>%
                             mutate(num_weeks = n()) %>%
                           ungroup() %>%
                           dplyr::filter(num_weeks > 0) %>%
         dplyr::mutate(event_recat = ifelse(event_cat1 == 0 | is.na(event_cat1), 'Week with No Suicidal Event', 'Week With Suicidal Event')), 
         aes(x = week_num, y = factor(ID), fill = event_recat)) +
  geom_tile(color = 'black') +
  scale_fill_manual(values = c('gray50', 'red')) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_x_continuous(limits = c(0,27)) +
  labs(x = 'Number of Weeks Following Baseline Visit', y = 'Participants', fill = NULL, title = '')

ema_trigger_timing_plot = ggplot(data = df %>%
                             group_by(ID) %>%
                             mutate(num_weeks = n()) %>%
                           ungroup() %>%
                           dplyr::filter(num_weeks > 0) %>%
         dplyr::mutate(event_recat = case_when(
           ema_trigger == 0 ~ "Week With No Clinically Significant Suicide Ideation",
            ema_trigger == 1 ~ "Week With Clinically Significant Suicide Ideation",
            is.na(ema_trigger) ~ "Missing Weekly STB Survey")),
         aes(x = week_num, y = factor(ID), fill = event_recat)) +
  geom_tile(color = 'black') +
  scale_fill_manual(values = c('white', 'red', 'gray50')) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  scale_x_continuous(limits = c(0,27)) +
  labs(x = 'Number of Weeks Following Baseline Visit', y = 'Participants', fill = NULL, title = '')

event_occur_grid = cowplot::plot_grid(event_timing_plot, ema_trigger_timing_plot, nrow = 2,
                                      align = 'v', axis = 'lr', 
                                      labels = c('A: Suicidal Events', 'B: Clinically Significant Suicide Ideation'))

cowplot::save_plot(event_occur_grid, file = 'plots/supplement/event_timing_descriptive.png', base_height = 8, base_width = 10)



timing_plot = cowplot::plot_grid(week_timeline_plot, event_predict_geolocation_plot2, labels = c('A', 'B'),
                   rel_widths = c(1.1, 1), align = 'h', axis = 'bt')

cowplot::save_plot(timing_plot, file = 'plots/fig2_timing_plot_geo_events_event_sensitivity.png', base_height = 4, base_width = 10)

```

##### Supplemental Table For Timing for Events Analysis
```{r}
tidy_weeks %>% mutate(aOR = round(Estimate, 2), 
         `95% CI`=paste0(round(Q2.5,2), ', ', round(Q97.5, 2))) %>%
  dplyr::select(timing, aOR, `95% CI`) %>%
  write.csv(., file = 'tables/homestay_timing_table_event_sensitivity.csv', row.names = FALSE)


tidy_event_predict_geo %>%
  mutate(`95% CI`=paste0(round(conf.low,2), ', ', round(conf.high, 2)),
         estimate = format(round(estimate, 2), nsmall=2),
         std.error=format(round(std.error,3), nsmall=3),
         p.value = format(round(p.value, 2), nsmall=2)) %>%
  dplyr::select(`Variable`=outcome, Beta=estimate, SE=std.error, `95% CI`, p=p.value) %>%
  write.csv(file = 'tables/suicide_events_predict_homestay_table_event_sensitivity.csv', row.names = FALSE)
```

### Time Since Baseline & Suicide Outcomes

```{r}
df %>%
  pivot_longer(c(home_h_dev, entropy_dev, travel_km_dev)) %>%
  ggplot(data= ., aes(x = week_num, y = value, color = name)) +
  geom_smooth(method = 'lm')
```

#### GPS variables as a function of week_num only
```{r}
entropy_time= lmerTest::lmer(data = df, entropy_dev ~ week_num + (week_num|ID))
homestay_time =lmerTest::lmer(data = df, home_h_dev ~ week_num + (week_num|ID))
travel_time= lmerTest::lmer(data = df, travel_km_dev ~ week_num + (week_num|ID))

sjPlot::tab_model(entropy_time, homestay_time, travel_time)
```

##### Model events as a function of week_num only
```{r}
fit_cat1_within_only_time = glmer(event_cat1  ~  week_num + (1|ID) , data = df,
              family = binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

fit_trigger_only_time = glmer(ema_trigger  ~  week_num + (1|ID) , data = df,
              family = binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

tab_model(fit_cat1_within_only_time, fit_trigger_only_time, show.stat = TRUE)

sjPlot::plot_model(fit_trigger_only_time, type = 'eff')
```


### Group differences / COVID

```{r}
load('/Volumes/columbia/MAPS_Data/Analysis/pab_maps_scripts/missing_data/cleaned_data/gps_data_missingness.rda')
load('/Volumes/columbia/MAPS_Data/Analysis/pab_maps_scripts/cleaned_data/covid.Rda')
covid = dplyr::select(covid, date, schoolclosure, stayathome, site) 
covid$schoolclosure = as.numeric(as.character(covid$schoolclosure))
covid = mutate(covid, site = ifelse(site=='PITT', 'UPMC', site))

daily = left_join(combined_gps, dplyr::select(selfreport_v8, ID, GROUP, Control, SITE, sex, age, iPhone, contains('SSI')) %>% mutate(ID = as.character(ID)), by = 'ID')

daily = dplyr::select(daily, date = dt_feature, site=SITE, everything())
daily = mutate(daily, site = ifelse(startsWith(x = ID, prefix = '1'), 'CUIMC', 'UPMC'))


daily = left_join(daily, covid, by = c('date', 'site'))
daily = mutate(daily, schoolclosure = ifelse(date > '2022-01-01', 0, schoolclosure),
                      schoolclosure = ifelse(date < '2020-03-01', 0, schoolclosure))


daily = daily %>%
  group_by(ID) %>%
  mutate(personal_time = date - min(date)) %>%
  dplyr::filter(personal_time < 200)

daily = mutate(daily, amt_home_day_h = amt_home_day_minutes/60,
               age = scale(age, center=TRUE, scale = FALSE))

homestay_group_schoolclosure = lmerTest::lmer(data = daily, amt_home_day_h ~ GROUP + schoolclosure + personal_time +  site + sex + age + iPhone + (1|ID))

homestay_ctrl_schoolclosure = lmerTest::lmer(data = daily, amt_home_day_h ~ Control + schoolclosure + personal_time + site + sex + age + iPhone + (1|ID))

homestay_ssi_schoolclosure = lmerTest::lmer(data = daily, amt_home_day_h ~ SSI19_baseline_imp1month + schoolclosure + personal_time + site + sex + age + iPhone + (1|ID))

tab_model(homestay_group_schoolclosure, homestay_ctrl_schoolclosure, homestay_ssi_schoolclosure)
```

