---
title: "Data Vis & Models"
author: "Ranqing Lan - updated Paul Bloom"
date: "Created 2023-04-21, updated 2023-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(lme4)
library(Matrix)
library(lmerTest)
library(brms)
library(rstan)
library(emmeans)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(crosstalk)
library(shiny)
library(DT)
library(readxl,quietly=TRUE,warn.conflicts=FALSE)
library(lubridate,quietly=TRUE,warn.conflicts=FALSE )
library(zoo)
library(psych)
library(openxlsx)
library(corrplot)
library(sjmisc)
library(sjlabelled)
library(sjPlot)
#library(coefplot2)
library(glmmTMB)
library(gridExtra)
library(grid)
library(ggplot2)
options(scipen = 999)
```

# Pull Data In & Prep

```{r, include = FALSE}
df <- read_csv('cleaned_data/weekly_data_with_stb.csv') %>%
  mutate(home_h_idmean = home_min_idmean/60)
df_updated_homestay= read_csv('cleaned_data/weekly_data_with_stb_multihome.csv') %>%
  mutate(home_h_idmean = home_min_idmean/60) %>%
  dplyr::select(ID, week_start_wednesday, home_h_dev_either=home_h_dev, home_h_idmean_either = home_h_idmean)


df = left_join(df, df_updated_homestay, by = c('ID', 'week_start_wednesday'))

# week # check for schoolclosure
df %>%
  mutate(schoolclosure = dplyr::recode(schoolclosure, '1'='Closed (Remote Only)', '0'='Open (In-Person)'),
         site = ifelse(startsWith(as.character(ID), '1'), 'New York City', 'Pittsburgh')) %>%
  ggplot(data=., aes(x = week_start_wednesday, y = as.character(ID), fill = schoolclosure)) +
  geom_tile() +
  facet_grid(rows = vars(site), drop = TRUE, scales = 'free_y') +
  theme_bw() +
  theme(axis.text.y = element_blank()) +
  labs(x = 'Date', y = 'Participants', fill = 'Local Public High School Status')


## Grand mean scale geolocation features
df <- df %>%
  group_by(ID) %>%
  mutate(
    entropy_dev = scale(entropy_dev, center = FALSE, scale = TRUE),
    home_h_dev = scale(home_h_dev, center = FALSE, scale = TRUE),
    home_h_dev_either = scale(home_h_dev_either, center = FALSE, scale = TRUE),
    travel_km_dev = scale(travel_km_dev, center = FALSE, scale = TRUE),
  )
```

## Results {.tabset}

### Within-participant GPS variable correlations

```{r, include=FALSE}
within_cors = df %>%
  group_by(ID) %>%
  summarise(`Travel Distance~Entropy` = cor(travel_km_dev, entropy_dev, use = 'pairwise.complete.obs'),
            `Travel Distance~Homestay (EARS)` = cor(travel_km_dev, home_h_dev, use = 'pairwise.complete.obs'),
            `Entropy~Homestay (EARS)` = cor(entropy_dev, home_h_dev, use = 'pairwise.complete.obs'),
            `Travel Distance~Homestay (EARS+)` = cor(travel_km_dev, home_h_dev_either, use = 'pairwise.complete.obs'),
            `Entropy~Homestay (EARS+)` = cor(entropy_dev, home_h_dev_either, use = 'pairwise.complete.obs'),
            `Homestay (EARS)~Homestay (EARS+)` = cor(home_h_dev, home_h_dev_either, use = 'pairwise.complete.obs'))
```

#### Boxplot to visualize correlation distributions

```{r}
within_cor_plot = within_cors %>%
  pivot_longer(contains('~')) %>%
  ggplot(data = ., aes(x = name, y = value)) +
  geom_hline(yintercept = 0, lty = 2, color = 'purple') +
  geom_boxplot() +
  coord_flip() + 
  labs(y = 'Distribution of Within-Participant Correlations\nCalculated on Weekly Metrics', x= NULL) +
  theme_bw()

ggsave(within_cor_plot,
       width = 7, height = 6,
       file = 'plots/supplement/gps_measures_within_participant_correlations.png')
```

#### Means and standard deviations across participants in correlations

```{r}
within_cors %>%
  pivot_longer(contains('~')) %>%
  group_by(name) %>%
  summarise(mean_correlation = mean(value, na.rm = TRUE),
            sd_correlation = sd(value, na.rm=TRUE))
```

### Histograms


#### Within-participant variables

```{r}
entropy_hist <- ggplot(df, aes(x= entropy_dev)) +
  geom_histogram( color = "black", fill='red', alpha=1, position='identity') + theme_classic() + 
  labs(x="Mean Daily Entropy\nAveraged Weekly & Within-Participant Mean-Centered", y = "Count", color = "black")


# for homestay, 2 rows with histograms for the original EARS home location calculation + updated calculation
homestay_hist <- ggplot(df %>%
                              pivot_longer(contains('home_h_dev')) %>%
                              dplyr::filter(!grepl('week', name)) %>%
                          dplyr::mutate(name = recode(name, 'home_h_dev'='EARS', 'home_h_dev_either'='EARS+')),
                             aes(x= value)) +
  geom_histogram( color = "black", fill='gray', alpha=1, position='identity') + theme_classic() + 
  stat_summary(aes(x = 0.1, y = value, xintercept = stat(y)), 
               fun.y = mean, geom = "vline", lwd = 1, color = 'purple') +
  labs(x="Mean Daily Homestay Hours\nAveraged Weekly & Within-Participant Mean-Centered", y = "Count", color = "black") +
  facet_grid(rows = vars(name))


travel_km_hist <- ggplot(df, aes(x= travel_km_dev)) +
  geom_histogram( color = "black", fill='lightblue', alpha=1, position='identity') + theme_classic() +
  labs(x="Mean Daily Distance (km) Traveled\nAveraged Weekly & Within-Participant Mean-Centered", y = "Count", color = "black")
```

#### Between-participant variables (participant means)

```{r}
# Just 1 row needed per person since we are looking at person-level averages
btw_df = df %>%
  group_by(ID) %>%
  filter(row_number()==1)

entropy_hist_mean <- ggplot(btw_df, aes(x= entropy_idmean)) +
  geom_histogram( color = "black", fill='red', alpha=1, position='identity') + theme_classic() + 
  labs(x="Mean Daily Entropy\nParticipant-Level Means", y = "Count", color = "black")

# for homestay, 2 rows with histograms for the original EARS home location calculation + updated calculation
homestay_mean_hist <- ggplot(btw_df %>%
                              pivot_longer(contains('home_h_idmean')) %>%
                              dplyr::filter(!grepl('week', name)) %>%
                          dplyr::mutate(name = recode(name, 'home_h_idmean'='EARS', 'home_h_idmean_either'='EARS+')),
                             aes(x= value)) +
  geom_histogram( color = "black", fill='gray', alpha=1, position='identity') + theme_classic() + 
  stat_summary(aes(x = 0.1, y = value, xintercept = stat(y)), 
               fun.y = mean, geom = "vline", lwd = 1, color = 'purple') +
  labs(x="Mean Daily Homestay Hours\nParticipant-Level Means", y = "Count", color = "black") +
  facet_grid(rows = vars(name))

travel_km_mean_hist <- ggplot(btw_df, aes(x= travel_km_mean)) +
  geom_histogram( color = "black", fill='lightblue', alpha=1, position='identity') + theme_classic() +
  labs(x="Mean Daily Distance (km) Traveled\nParticipant-Level Means", y = "Count", color = "black")

```


#### Comparison of EARS vs EARS+ homestay
```{r}
btw_df = mutate(btw_df, home_h_diff = home_h_idmean_either - home_h_idmean)
m_ears_comp = lm(data = btw_df, home_h_diff ~ 1)
tab_model(m_ears_comp, show.stat = TRUE, show.df = TRUE)
summary(btw_df$home_h_idmean)
sd(btw_df$home_h_idmean, na.rm=TRUE)

summary(btw_df$home_h_idmean_either)
sd(btw_df$home_h_idmean_either, na.rm=TRUE)
```

### 3. Missingness Plots


```{r}
missingness_gps <- df %>%
  select(ID, week_num, gps_missing3) %>%
  group_by(ID) %>%
  mutate(data_collected0 = ifelse(gps_missing3==1, 0,1),
         data_collected = as.factor(data_collected0),
         sum_datacollected = sum(data_collected0, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(ID, desc(sum_datacollected)) %>%
  mutate(data_collected = dplyr::recode(data_collected, '0' ='\u22642 Days With GPS Data Observed', '1'='\u22653 Days with GPS Data Observed')) %>% 
  ggplot(data = ., aes(x = week_num, y = fct_reorder(.f = as.factor(ID), .x = sum_datacollected))) +
  geom_tile(aes(fill = data_collected)) + 
  labs(y = 'Participants', x = "Number of Weeks Following Baseline Visit", color = "Data Collected", color = "black", fill = NULL) +
  ggtitle("A: Missingness of GPS Data") +
  theme_linedraw() +
  theme(axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = 'bottom') +
  scale_x_continuous(expand = c(0, 0), limits = c(0,26))

missingness_ema_trigger <- df %>%
  select(ID, week_num, ema_trigger) %>%
  group_by(ID) %>%
  mutate(data_collected0 = ifelse(is.na(ema_trigger), 0,1),
         data_collected = as.factor(data_collected0),
         sum_datacollected = sum(data_collected0, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(ID, desc(sum_datacollected)) %>%
  mutate(data_collected = dplyr::recode(data_collected, '0' ='Missing', '1'='Observed')) %>% 
  ggplot(data = ., aes(x = week_num, y = fct_reorder(.f = as.factor(ID), .x = sum_datacollected))) +
  geom_tile(aes(fill = data_collected)) + 
  labs(y = 'Participants', x = "Number of Weeks Following Baseline Visit", color = "Data Collected", color = "black", fill = NULL) +
  ggtitle("B: Missingness of Weekly STB Surveys") +
  theme_linedraw() +
  theme(axis.text.y=element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = 'bottom') +
  scale_x_continuous(expand = c(0, 0), limits = c(0,26))
```


```{r}
gps_distributions_within = cowplot::plot_grid(entropy_hist, homestay_hist, travel_km_hist,
                                       align = 'h', axis = 'bt',
                   labels = c('A', 'B', 'C'))

gps_distributions_within

gps_distributions_btw = cowplot::plot_grid(entropy_hist_mean, homestay_mean_hist, travel_km_mean_hist,
                                       align = 'h', axis = 'bt',
                   labels = c('A', 'B', 'C'))

gps_distributions_btw

missingness_grid = cowplot::plot_grid(missingness_gps, missingness_ema_trigger)
missingness_grid
```


#### Save plot files
```{r}
cowplot::save_plot(gps_distributions_within, filename = 'plots/supplement/gps_within_plot.png',
                   base_width = 9, base_height = 5)


cowplot::save_plot(gps_distributions_btw, filename = 'plots/supplement/gps_between_plot.png',
                   base_width = 9, base_height = 5)


cowplot::save_plot(missingness_grid, filename = 'plots/supplement/missing_data.png',
                   base_width = 11, base_height = 4)
```